---
title: "Exercises 1"
author: "Bas Machielsen"
date: '2022-07-11'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Exercise 1 - Logistic Regression

Use the Boston data (e.g. from the package MASS. Create a dependent dummy variable for house price which is is higher than 200K.

```{r}
library(MASS); library(tidyverse)
boston <- MASS::Boston

boston <- boston %>%
  mutate(dummy = if_else(medv > 20, 1, 0))
```

Split the data (use set.seed(55)) into a train and a test set.

- I do this through Python

```{python}
import random
import numpy as np
from sklearn.model_selection import train_test_split
import statsmodels.api as sm
import pandas as pd

r.boston.iloc[:,0:13]

X_train, X_test, y_train, y_test = train_test_split(
  r.boston.iloc[:,0:13], r.boston.iloc[:,14], test_size=0.5, 
  random_state=55)
  
```

Based on the other variables in the data, predict whether a house would be priced over 200K or below.

```{python}

clf = sm.Logit(y_train, X_train).fit()

clf.summary()

```    


Create a confusion matrix to evaluate the accuracy

```{python}
def confusion_matrix(y_test,x_test, logit_model, threshold=0.5):
  
  predictions = logit_model.predict(x_test)
  
  data = (pd.DataFrame(np.array([y_test, predictions]).
  transpose(),
  columns=['real','predict']))
  
  data['predict'] = data['predict'] > threshold
  
  predicted_no_actual_yes, predicted_yes_actual_yes = (
    data.groupby('predict')['real'].apply(lambda x:
      (x==1).sum()).reset_index(name='count').iloc[:,1])
  predicted_no_actual_no, predicted_yes_actual_no = (
    data.groupby('predict')['real'].apply(lambda x:
      (x==0).sum()).reset_index(name='count').iloc[:,1] )
  
  
  out = np.array([[predicted_yes_actual_yes, predicted_no_actual_yes],
  [predicted_yes_actual_no, predicted_no_actual_no]])
  
  b_out = (pd.DataFrame(out, index=['actual_yes', 'actual_no'],
  columns=['predicted_yes', 'predicted_no']))
  return b_out
  
```


```{python}
confusion_matrix(y_test, X_test, clf)
```

## Activation Functions

- What is the role of the activation function?

To scale down the predicted values towards a reasonable range (e.g. $(0,1)$). 

- Code yourself the sigmoid function, plot it (say for the [-5,5] range)

- Code yourself the relu function, plot it.

- Code yourself the leaky relu function, plot it.

- Code yourself the swish function, plot it.

- Compute and plot the derivatives of those functions2.

- Recreate the softmax chart presented during the lecture.


