---
title: "Practical sessions"
author: ""
date: ''
output: 
  html_document:
    toc: true
    number_sections: false
    theme: simplex
    highlight: espresso
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# for generic R help and tutorials
# help.start()

#  Setup  --------------------
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# install.packages("magrittr")
# install.packages("psych")
library(magrittr)
library(MASS)
library(psych)

```

## Gradient Descent in action

(a) Simulate some data:
```{r chunk2, echo= T, tidy=F}
x <- runif(1000, -5, 5)
y <- x + rnorm(1000) + 3
```

(b) Create a squared root cost function.

(c) Set iteration number to 100 and the learning rate to 0.01. Initialize the intercept
and sloe to zero.

(d) Loop through iterations and nudge the parameters based on the gradient.

(e) Keep the history and plot the cost function over time.


## First neural network

(a) Load the Boston data and split it using `set.seed(55)` again.

(b) Scale the data and create a confusion matrix based on the standard logistic
regression.

(c) Create predictions (remember to scale also the test set). Report the accuracy.

(d) Use the `neuralnet` package to create a simple neural network with 1 hidden
layer, `backprop` algorithm, SSE cost function and a 0.1 learning rate parameter.

(e) Report the accuracy of this network after predicting the test set.

(f) Now change the network architecture to c(8,2), which means 2 hidden layers
with 8 and 2 neurons respectively. Is accuracy increased?

(g) Apply early stopping and check the accuracy again.

(h) Plot model architecture.

## MNIST data

The MNIST is a database of handwritten digits, and is commonly used for training various image processing systems. The training set contains 60K examples, and the text set contains 10K examples.

(a) Load the data.

(b) Create a `min_max_normalization` function, and apply it on the data.

(c) Plot the few of the images^[You can, but you don’t have to, use the function provided in the Appendix].

(d) Use the `nnet` package and perform multinomial regression.

(e) Use the predict function to predict the test data.

(f) Create a confusion matrix^[You can use the `caret` package for example], and compute the accuracy. We will later compare it
to our neural network model.

(g) Create a neural network model in Keras. Choose yourself the number of neurons,
batch size, and the number of epochs. Use first a gradient descent optimizer.

(h) Keep all your chosen settings, but use the ADAM optimizer now.


##### Footnotes {#endnotes}

<div id="refs"></div>

#### Appendix

```{r chunk3, echo= T, tidy=F}
Show_label <- function(row, data) {
  tmp <- data.frame(
    x = rep(1:28, times = 28),
    y = rep(28:1, each = 28),
    shade = as.numeric(data[row, -1])
  )
  ggplot(data = tmp) +
    geom_point(aes(x = x, y = y, color = shade),
               size = 11,
               shape = 15) + theme(
                 waxis.line =
                   element_blank(),
                 axis.text.x = element_blank(),
                 axis.text.y = element_blank(),
                 axis.ticks = element_blank(),
                 axis.title.x = element_blank(),
                 axis.title.y = element_blank(),
                 legend.position = "none",
                 panel.background = element_blank(),
                 panel.border = element_blank(),
                 panel.grid.major = element_blank(),
                 panel.grid.minor = element_blank(),
                 plot.background = element_blank()
               ) +
    scale_color_gradient(low = "white", high = "black") +
    geom_text(aes(x = 28, y = 28), label = data[row, 1])
}
```


```{js jschunk, echo=FALSE}
$(document).ready(function() {
  $('.footnotes ol').appendTo('#endnotes');
  $('.footnotes').remove();
});
```


